The Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals

### 2022-04-19 ms/ms.Rmd
update ms

---
title: "The Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals"
author: "F. Débarre, E. Lecoeur, L. Guimier, M. Jauffret-Roustide, A.-S. Jannot"
output:
  pdf_document: 
    keep_tex: yes
    latex_engine: pdflatex
    includes:
      in_header: template.tex
  bookdown::html_document2:
    self_contained: no
    keep_md: true
    number_sections: no
  bookdown::word_document2: 
    number_sections: no
bibliography: VaccinationInequalities.bib
---

@@

## Data sources
-  Assurance Maladie: <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-devaccination-par-epci/> <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-de-vaccination-parcommune/information/>
-  INSEE: <https://www.insee.fr/fr/statistiques/5359146#consulter>

@@

All analysis code is available at XXX; analyses were done in `r R.Version()$version.string`.

>>>>

---
title: "The Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals"
author: "F. Débarre, E. Lecoeur, L. Guimier, M. Jauffret-Roustide, A.-S. Jannot"
output:
  bookdown::pdf_document2:
    keep_tex: yes
    latex_engine: pdflatex
    includes:
      in_header: template.tex
    fig_caption: yes
    number_sections: no
    toc: no
  bookdown::html_document2:
    self_contained: no
    keep_md: true
    number_sections: no
  bookdown::word_document2: 
    number_sections: no
bibliography: VaccinationInequalities.bib
---

@@

## Data sources
-  Assurance Maladie: <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-devaccination-par-epci/> <https://datavaccin-covid.ameli.fr/explore/dataset/donnees-de-vaccination-parcommune/information/>
-  INSEE: <https://www.insee.fr/fr/statistiques/5359146#consulter>
-  2017 Presidential election: <https://www.data.gouv.fr/fr/datasets/election-presidentielle-des-23-avril-et-7-mai-2017-resultats-definitifs-du-1er-tour-par-communes/#resource-d282e53a-d273-425d-95bb-8a0d7632c79a-header>

@@

All analysis code is available at \url{https://github.com/flodebarre/covid-passports-france} (and will be stored on a permanent repository upon acceptance); analyses were done in `r R.Version()$version.string`.

### 2022-04-19 ms/ms.tex
update ms

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals},
  pdfauthor={F. Débarre, E. Lecoeur, L. Guimier, M. Jauffret-Roustide, A.-S. Jannot},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}
\usepackage{lineno} % line numbering
\linenumbers

@@

\begin{itemize}
\tightlist
\item
  Assurance Maladie:
  \url{https://datavaccin-covid.ameli.fr/explore/dataset/donnees-devaccination-par-epci/}
  \url{https://datavaccin-covid.ameli.fr/explore/dataset/donnees-de-vaccination-parcommune/information/}
\item
  INSEE: \url{https://www.insee.fr/fr/statistiques/5359146\#consulter}
\end{itemize}

@@

All analysis code is available at XXX; analyses were done in R version
4.0.4 (2021-02-15).

@@

We investigated the associations between each of the 141 indicators
characterizing districts of residence, and the fact of having received
at least one Covid-19 vaccine dose, on the whole population of mainland
France. Two indicators were among the top five most associated one at
all time points (see Figure @ref(fig:figManhattan)): the share of local
income coming from unemployment benefits (\texttt{Unemployment\_Benef};
strongest association on 2022-01-23, \(OR = 0.716\)) and vote for the
``anti-establishment'' political party represented by the candidate
Asselineau (\texttt{Asselineau}; \(OR = 0.712\) on 2022-02-20). The
three other most associated indicators did not change in the later dates
that we considered, and were the proportion of immigrants in the
district (\texttt{Immigrant}; \(OR = 0.713\) on 2022-02-20), the
district's relative position along a North-West--South-East gradient
(\texttt{NO-SE}; \(OR = 0.745\) on 2021-12-12) and the proportion of
overcrowded households (\texttt{Overcrowding\_rate}; \(OR = 0.738\) on
2022-01-23).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figManhattan-1.pdf}
\caption{Manhattan plots of the Odds ratios for each of the indicator of
our dataset, by date. Left column: around the Sanitary Pass
implementation; right column: around the Vaccine Pass implementation.
The top odds ratios are labelled at each time point; the symbol next to
the name indicates the direction of the effect. The gray rectangle
corresponds to the 99\% percentile of odds ratios in the permuted data;
points falling in the rectangle are considered as non-significant.}
\end{figure}

Our odds ratio calculations were based on a crude version of each
indicator, which were dichotomized into values above or below the median
of each indicator. To better visualize the effects (or lack thereof) of
the sanitary and vaccine passes on vaccination rates over time, we
computed age-adjusted vaccination rates over time, by decile of three of
the most associated indicators, treated as factors (see Figure
@ref(fig:figOverTime)). The Sanitary Pass, implemented in the Summer
2021, led to an overall increase in vaccination rates; on the other
hand, the Vaccine Pass, implemented in the end of 2021, did not affect
the evolution of vaccination rates. Heterogeneities in vaccination rates
persisted after both types of pass; vaccination rates gradually decrease
by decile of each indicator, confirming the association of these
indicators with vaccination rates without threshold effect. Of note, for
the Unemployment and Asselineau vote indicators, the difference between
the 9th and 10th deciles appears to be much larger than between the
other consecutive deciles.

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figOverTime-1.pdf}
\caption{Age-adjusted vaccination rates among adults, over time, by
decile of each indicator (presented by a color gradient). The vertical
lines indicate the dates of announcements and implementations of the
sanitary and vaccine passes.}
\end{figure}

Finally, historically under-vaccinated areas in France stand out as
being less vaccinated against Covid-19, in particular the South-East
region (see Figure @ref(fig:figMap)).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figMap-1.pdf}
\caption{Adult vaccination rates by district of mainland France}
\end{figure}

>>>>

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={The Covid-19 vaccination policy did not solve vaccination inequities: a nationwide longitudinal study on 64.5 million individuals},
  pdfauthor={F. Débarre, E. Lecoeur, L. Guimier, M. Jauffret-Roustide, A.-S. Jannot},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{setspace}
\usepackage{lineno} % line numbering
\linenumbers

@@

\begin{itemize}
\tightlist
\item
  Assurance Maladie: \url{https://datavaccin-covid.ameli.fr/explore/dataset/donnees-devaccination-par-epci/} \url{https://datavaccin-covid.ameli.fr/explore/dataset/donnees-de-vaccination-parcommune/information/}
\item
  INSEE: \url{https://www.insee.fr/fr/statistiques/5359146\#consulter}
\item
  2017 Presidential election: \url{https://www.data.gouv.fr/fr/datasets/election-presidentielle-des-23-avril-et-7-mai-2017-resultats-definitifs-du-1er-tour-par-communes/\#resource-d282e53a-d273-425d-95bb-8a0d7632c79a-header}
\end{itemize}

@@

All analysis code is available at \url{https://github.com/flodebarre/covid-passports-france} (and will be stored on a permanent repository upon acceptance); analyses were done in R version 4.0.4 (2021-02-15).

@@

We investigated the associations between each of the 141 indicators characterizing districts of residence, and the fact of having received at least one Covid-19 vaccine dose, on the whole population of mainland France. Two indicators were among the top five most associated one at all time points (see Figure \ref{fig:figManhattan}): the share of local income coming from unemployment benefits (\texttt{Unemployment\_Benef}; strongest association on 2022-01-23, \(OR = 0.716\))
and vote for the ``anti-establishment'' political party represented by the candidate Asselineau (\texttt{Asselineau}; \(OR = 0.712\) on 2022-02-20).
The three other most associated indicators did not change in the later dates that we considered, and were the proportion of immigrants in the district (\texttt{Immigrant}; \(OR = 0.713\) on 2022-02-20),
the district's relative position along a North-West--South-East gradient (\texttt{NO-SE}; \(OR = 0.745\) on 2021-12-12)
and the proportion of overcrowded households (\texttt{Overcrowding\_rate}; \(OR = 0.738\) on 2022-01-23).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figManhattan-1.pdf}
\caption{\label{fig:figManhattan}Manhattan plots of the Odds ratios for each of the indicator of our dataset, by date. Left column: around the Sanitary Pass implementation; right column: around the Vaccine Pass implementation. The top odds ratios are labelled at each time point; the symbol next to the name indicates the direction of the effect. The gray rectangle corresponds to the 99\% percentile of odds ratios in the permuted data; points falling in the rectangle are considered as non-significant.}
\end{figure}

Our odds ratio calculations were based on a crude version of each indicator, which were dichotomized into values above or below the median of each indicator. To better visualize the effects (or lack thereof) of the sanitary and vaccine passes on vaccination rates over time, we computed age-adjusted vaccination rates over time, by decile of three of the most associated indicators, treated as factors (see Figure \ref{fig:figOverTime}). The Sanitary Pass, implemented in the Summer 2021, led to an overall increase in vaccination rates; on the other hand, the Vaccine Pass, implemented in the end of 2021, did not affect the evolution of vaccination rates. Heterogeneities in vaccination rates persisted after both types of pass; vaccination rates gradually decrease by decile of each indicator, confirming the association of these indicators with vaccination rates without threshold effect. Of note, for the Unemployment and Asselineau vote indicators, the difference between the 9th and 10th deciles appears to be much larger than between the other consecutive deciles.

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figOverTime-1.pdf}
\caption{\label{fig:figOverTime}Age-adjusted vaccination rates among adults, over time, by decile of each indicator (presented by a color gradient). The vertical lines indicate the dates of announcements and implementations of the sanitary and vaccine passes.}
\end{figure}

Finally, historically under-vaccinated areas in France stand out as being less vaccinated against Covid-19, in particular the South-East region (see Figure \ref{fig:figMap}).

\begin{figure}
\centering
\includegraphics{ms_files/figure-latex/figMap-1.pdf}
\caption{\label{fig:figMap}Adult vaccination rates by district of mainland France}
\end{figure}
